---
title: "Conditional & Combined Plots"
---

```{r, echo = FALSE}
diabetes <- readRDS(here::here("data", "diabetes.rds"))
titanic <- readRDS(here::here("data", "titanic.rds"))
```

We can convert most univariate visualizations that we create with Base R graphics into conditional plots by specifying a
grouping factor. We then get a single figure with group-specific versions of the univariate visualization.

### Boxplots

We can use conditional boxplots to compare the distribution of a variable accross groups. To create this type of figure,
we employ the so-called *formula interface* when calling the `boxplot()` function. In the following code, we use the
`diabetes` dataset to create sex-specific boxplots of disease progression.

```{r}
boxplot(progress ~ sex, data = diabetes)
```

When we use the formula interface, we define the plot's structure through an R formula object that we specify as the
first argument to `boxplot()`. In this formula, `progress ~ sex1`, we specify the plotted variable on the lefthand side
of the formula and we put the grouping factor on the righthand side of the formula. We also have to use the `data`
argument to specify a data frame where `boxplot()` can find the variables we're referencing in the formula.

```{webr-r}
#| autorun: true
#| context: setup

download.file("https://openstatprog.com/raw-data/titanic.rds", "titanic.rds")
titanic <- readRDS("titanic.rds")
```

::: practice

## Practice

Use the `titanic` dataset to create a conditional boxplot to visualize the conditional distributions of `age` within the
groups defined by the `survived` variable.

- What does this figure tell you about the ages of survivors (`survived` = 1) vs. non-survivors (`survived` = 0)?

*NOTE:* The `titanic` dataset is already loaded into the working directory of this *webr* session.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}

```

## {{< var tab_title.solution >}}
```{r}
boxplot(age ~ survived, data = titanic)
```

The median age does not seem to differ much between the two groups, but the survivors' distribution appears to be more
symmetric and there are fewer old survivors. The non-survivors' distibution is positively skewed by the
disproportionately higher number of fatalities among older age groups and relatively fewer fatalities among young
passangers.
::::

:::

### Bar Plots

Conditional bar plots visualize two dimensional contingency tables. So, these types of figure allow us to evaluate how
the distribution of one categorical variable differs across levels of some other variable.

To create a conditional bar plot, we fist create a two-way contingency table, and then we plot that table with the
`barplot()` funciton. In the following code we use the `titanic` dataset to visualize the conditional distributions of
ticket classes for the two survival groups.

```{r}
# Cross-tabulate ticket class and survival status
tab <- table(titanic$class, titanic$survived)

# Visualize the table created above
barplot(tab)
```

The default barplot created above suffers from two important problems.

1. The bars for each ticket class group are stacked, so the meaning of the values shown on the y-axis is ambiguous
1. We don't know which ticket group is represented by which color.

We can solve these problems by plotting the bars beside each other by setting `beside = TRUE` and drawing a legend by
setting `legend = TRUE`.

```{r}
# Create a better visualization of the table
barplot(tab, beside = TRUE, legend = TRUE)
```

### Histograms

When generating histogams, we can't really create the same kind of conditional plots that we can with boxplots and bar
plots, but we can overlay different plot layers to add useful information to our figure.

```{r}
# Create a basic histogram of blood glucose levels
hist(diabetes$glu, probability = TRUE)

# Calculate the kernel density estimate of blood glucose levels
d <- density(diabetes$glu)

# Overlay the kernel density curve on the histogram
lines(y = d$y, x = d$x, col = "red")
```

```{r}
# Create a basic histogram of average blood pressure levels
hist(diabetes$bp)

# Calculate the limits of the central 95% of observations
ci <- quantile(diabetes$bp, prob = c(0.025, 0.975))

# Add dashed lines to denote the 5% tail area cutoffs
abline(v = ci, lty = 2, lwd = 2)
```

```{r}
# Create a scatterplot of blood pressure against blood glucose
plot(bp ~ glu, data = diabetes, pch = 19)

# Estimate the linear regression of blood pressure on blood glucose
lmOut <- lm(bp ~ glu, data = diabetes)

# Add the best fit line to the scatterplot
abline(reg = lmOut, lwd = 2, col = "blue")
```
