---
title: "Conditional Plots"
---

```{r, echo = FALSE}
diabetes <- readRDS(here::here("data", "diabetes.rds"))
titanic <- readRDS(here::here("data", "titanic.rds"))
```

We can convert most univariate visualizations that we create with Base R graphics into conditional plots by specifying a
grouping factor. We then get a single figure with group-specific versions of the univariate visualization.

### Boxplots

We can use conditional boxplots to compare the distribution of a variable accross groups. To create this type of figure,
we employ the so-called *formula interface* when calling the `boxplot()` function. In the following code, we use the
`diabetes` dataset to create sex-specific boxplots of disease progression.

```{r}
boxplot(progress ~ sex, data = diabetes)
```

When we use the formula interface, we specify the plotted variable on the left hand side of the formula and we put the
grouping factor on the righthand side of the formula. We also have to specify a data frame where `boxplot()` can find
the variables we're referencing in the formula.

```{webr-r}
#| autorun: true
#| context: setup
#| 
download.file("https://raw.githubusercontent.com/p3piit/exp_rawdata/main/diabetes.rds", "titanic.rds")
titanic <- readRDS("titanic.rds")
library(magrittr)
```

::: practice

## Practice

Use base R graphics and the `titanic` data to create conditional boxplots, where plots of `age` are conditioned on
`survived`. What does this figure tell you about the ages of survivors (`survived` = 1) vs. non-survivors (`survived` =
0)?

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}

```

## {{< var tab_title.solution >}}

As we can see, the age distributions of survivors and non-survivors are quite similar. Both groups have nearly the same
median age and a comparable spread of ages, indicating that age alone does not appear to be a strong distinguishing
factor between survival and non-survival in this dataset. 

```{r}
boxplot(age ~ survived, data = titanic)
```
::::

:::
