---
title: "Conclusion"
format:
  html: 
    toc: false
    link-external-icon: false
    link-external-newwindow: false
---

```{r include = FALSE}
source(here::here("R", "buttons.R"))
```

By completing this module, you've taken your first steps toward data visualization with R. You've learned how to use
Base R graphics to create some of the most useful types of data visualization: histograms, boxplots, bar plots, and
scatterplots. You also learned how to convert univariate plots into conditional plots and how to add layers to plots to
increase the information provided by a single figure.

## Next Steps

Base R graphics are a solid starting point, and the kinds of figures we covered in this module are a useful for model
diagnostics, data screening, and exploratory data analysis. To create complex, publication-ready figures, however, you
will probably want something more. To continue your data visualization journey, we recommend following the *Data
Visualization with GGPlot* module where you will learn how to use the GGPlot visualization suite to create
publication-quality figures in R.

```{r}
#| echo: false
#| results: asis
# next_module_button("https://www.openstatprog.com/ggplot/")
next_module_button("in_progress.html")
```
