---
title: "Basic Visualization"
page-navigation: false
format:
  html: 
    toc: false
---

```{r, echo = FALSE, include = FALSE}
source(here::here("R", "buttons.R"))
```

Welcome to the module on basic data visualization in R. Graphical visualizations are a crucial tool in data analysis
because they allow us to explore, summarize, and communicate patterns in data that would be difficult to detect through
numbers alone. Well-chosen visualizations can reveal trends, clusters, correlations, and outliers at a glance, helping
to guide further analysis and support clear, persuasive communication of results.

In this module, we will explore the main plotting capabilities built into base R, focusing on four foundational chart
types: scatterplots, histograms, boxplots, and density plots. Each section combines conceptual explanation with hands-on
examples to show how to create and customize these plots in R, highlighting how they can be adapted to different
datasets and analysis goals.

## Learning Goals

After completing this module, you will be able to use Base R functions to create:

1. Scatterplots to investigate relationships between two variables
2. Histograms to examine the distribution of a numeric variable
3. Boxplots to summarize distributions, detect outliers, and compare groups
4. Density plots to visualize approximations of a variable’s distribution

Click the button below to get started with the first tutorial.

```{r}
#| echo: false
#| results: asis
begin_button("/tutorials/basic_graph/index.html")
```
