---
title: "Histograms"
---

```{r, echo = FALSE}
diabetes <- readRDS(here::here("data", "diabetes.rds"))
```

Histograms are one of the most straightforward and useful ways to visualize the distribution of a numeric variable. A
histogram shows the distribution of a single numeric variable by dividing its range into bins (intervals), counting how
many observations fall into each bin, and visualizing these counts as bars in a plot. Histograms are meant to represent
a variables probability distribution, so they're useful for understanding the shape, spread, and skewness of numeric
varaibles.

The `hist()` function in Base R creates histograms. For example, the following code visualizes the distribution of blood
glucose levels in the diabetes dataset.

```{r}
hist(diabetes$glu)
```

Like the plot() function, `hist()` accepts many parameters to customize the appearance, such as `main` for the title,
`xlab` and `ylab` for axis labels, and `xlim` or `ylim` to set axis ranges. We can also control the number of bins with
the `breaks` argument:

```{r}
hist(diabetes$glu, main = "Broad bins", breaks = 5)   # Very broad bins
hist(diabetes$glu, main = "Small bins", breaks = 50)  # Many small bins
```

Styling options let us adjust colors and axis interpretation:

```{r}
hist(diabetes$glu,
     col = "pink",
     border = "blue",
     probability = TRUE)  # y-axis shows probability density
```

Setting `probability = TRUE` makes the histogram’s total area sum to 1, which is useful when comparing with density
plots.

```{webr-r}
#| autorun: true
#| context: setup
#| 
download.file("https://raw.githubusercontent.com/p3piit/exp_rawdata/main/diabetes.rds", "diabetes.rds")
diabetes <- readRDS("diabetes.rds")
library(magrittr)
```

::: practice

## Practice

Make an histogram showing the age distribution in the `diabetes` dataset, with small bins, green columns and an
appropiate title.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}

```

## {{< var tab_title.solution >}}


```{r}
hist(diabetes$age,
     col = "green",
     breaks = 50,
     main = "Age distribution")  

```
::::

:::
